% Choose one to switch between slides and handout
\documentclass[]{beamer}
%\documentclass[handout]{beamer}

% Video Meta Data
\title{Bitcoin, Blockchain and Cryptoassets}
\subtitle{Alternative Consensus Protocols}
\author{Prof. Dr. Fabian Schär}
\institute{University of Basel}

% Config File
\input{../config/config.tex}
\usetikzlibrary{calc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
	
	\thispagestyle{empty}
	\begin{frame}[noframenumbering]
		\titlepage
	\end{frame}
	
	
	%%%
	\begin{frame}{Why Consensus Matters}
		
		Distributed ledger as \color{focus} chain of transactions and states \color{black} whose \color{focus} compliance with an explicit rule set \color{black} is  attested by a reliable network of record keeping nodes.
		
		\uncover<2->{
			\vspace{1.5 em}
			\textbf{Account statement example:}
			
			\begin{center}
				\begin{tikzpicture}[scale=0.7, every node/.style ={scale=0.8}]
					\input{../assets/figures/account_statement_example}
				\end{tikzpicture}
			\end{center}
			
			\vspace{1 em}
			
			$\Rightarrow$ Value of the ledger content relies on the network attesting it.
		}
		
	\end{frame}
	%%%	
	
	%%%
	\begin{frame}{Measures Supporting Consensus}
		
		\textbf{Explicit and unambiguous rule set} for legitimate changes to the ledger and block sequence.
		\vspace{0.25 em}
		
		$\Rightarrow$ Invalid blocks are detected easily and unambiguously.
		\vspace{1.5 em}
		
		\textbf{Decision mechanism} for consensus over different, legitimate extensions of the ledger. 
		\vspace{0.25 em}
		
		$\Rightarrow$ Swiftly resolving situations of uncertainty.
		\vspace{1.5 em}
		
		\textbf{Incentive system} that rewards compliant behaviour and / or penalizes manipulation attempts.
		\vspace{0.25 em}
		
		$\Rightarrow$ Typically in native protocol asset, tying the participant's interest to the sustainable value of the network.
		
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{What Makes a Good Consensus Mechanism?}
		
		Suitability of a consensus mechanism \color{focus} depends on the purpose and user base \color {black} of a decentralized ledger.
		\vspace{1 em}
		
		\textbf{The Trilemma:}
		
		\begin{center}
			\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
				\input{../assets/figures/trilemma}
			\end{tikzpicture}
		\end{center}
		
		\textbf{Generalized Rule:} Subject to tadeoffs, i.e., not possible to achieve all three goals. 	
	\end{frame}
	%%%	
	
	%%%
	\begin{frame}{Popular Consensus Mechanisms}
		We will look at and compare the following consensus mechanisms:
		
		\vspace{2.5em}
		
		\centering
		\begin{tikzpicture}[scale=1, every node/.style={scale=1}]
			\input{../assets/figures/consensus_mechanisms}
		\end{tikzpicture}
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Proof of Work Trilemma}
		\begin{center}
			\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
				\input{../assets/figures/trilemma}
				\coordinate (AC) at ($(A)!0.4!(C)$);
				\node (pow) at ($(AC)!0.2!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (pow) [below = 0.2] {\texttt{PoW}};
			\end{tikzpicture}
		\end{center}
	
		\begin{description}[labelwidth=10em]
			\item[\textbf{Scalability}] Every node needs to process every transaction. Block creation is very resource intensive.
			\item[\textbf{Decentralization}] Open network with many participants. Mining pools compromise decentralization.
			\item[\textbf{Security}] Secured by ressource allocation. Simplicity increases security.
		\end{description}
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Proof of Stake}
		To participate in the consensus network, each node - called a validator - needs to deposit and lock protocol base assets. This is called staking.
		
		\vspace{1.5 em}
		\textbf{Block creation and transaction validation:}
		Block creators are selected at random in proportion to their stake. A set of other validators will then attest to the validity of the created blocks.
		
		\vspace{1.5 em}
		\textbf{Malicious and unresponsive validators:}
		Malicious behaviour is punished by slashing (destroying a significant part) of the staked assets of the offenders. Failure to participate forfeits any rewards and might be punished furhter.
		
		\vspace{1.5 em}
		\textbf{Incentive system:}
		Validators receive returns on their stake by performing their duties. These rewards are funded by transaction costs and/or newly generated assets (inflation).
		
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Proof of Stake Trilemma}
		\begin{center}
			\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
				\input{../assets/figures/trilemma}
				\coordinate (AC) at ($(A)!0.5!(C)$);
				\node (pos) at ($(AC)!0.5!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (pos) [below = 0.2] {\texttt{PoS}};
			\end{tikzpicture}
		\end{center}
		
		\begin{description}[labelwidth=10em]
			\item[\textbf{Scalability}] Only a set of validators need to process each transaction. Proposers are randomly selected.
			\item[\textbf{Decentralization}] Open network with many participants. Potential crowding out over time.
			\item[\textbf{Security}] Secured by staked capital. High complexity of the mechanism reduces security.
		\end{description}
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Proof of Authority}
		The consensus network consists of a \color{focus}small set of approved nodes\color{black} - called validators. These are often identified and have a reputation attached.
		
		\vspace{1.5 em}
		\textbf{Block creation and transaction validation:}
		Validators create and validate blocks. Other validators will attest to the validity of the created blocks.
		
		\vspace{1.5 em}
		\textbf{Malicious and unresponsive validators:}
		Malicious or unsresponsive behaviour is punished by excluding a validator and negatively impacting their reputation.
		
		\vspace{1.5 em}
		\textbf{Incentive system:}
		Block rewards are usually only the transaction costs.
		
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Proof of Authority Trilemma}
		\begin{center}
			\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
				\input{../assets/figures/trilemma}
				\coordinate (AC) at ($(A)!0.05!(C)$);
				\node (poa) at ($(AC)!0.95!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (poa) [below = 0.2] {\texttt{PoA}};
			\end{tikzpicture}
		\end{center}
		
		\begin{description}[labelwidth=10em]
			\item[\textbf{Scalability}] Very small set of validators and simple mechanism.
			\item[\textbf{Decentralization}] Closed network, validators can easily collude. Almost fully centralized.
			\item[\textbf{Security}] Simple mechanism and only approved validators provide a high security.
		\end{description}
	\end{frame}
	%%%
	
	%%%
	\begin{frame}{Consensus Trilemma}
		\begin{center}
			\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
				\input{../assets/figures/trilemma}
				
				\coordinate (AC) at ($(A)!0.4!(C)$);
				\node (pow) at ($(AC)!0.2!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (pow) [below = 0.2] {\texttt{PoW}};
				
				\coordinate (AC) at ($(A)!0.5!(C)$);
				\node (pos) at ($(AC)!0.5!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (pos) [below = 0.2] {\texttt{PoS}};
				
				\coordinate (AC) at ($(A)!0.05!(C)$);
				\node (poa) at ($(AC)!0.95!(B)$) [circle, fill = black, minimum width=0.2cm,minimum height=0.2cm]{};
				\node at (poa) [below = 0.2] {\texttt{PoA}};
			\end{tikzpicture}
		\end{center}
		
		\textbf{Conclusion:} All mechanisms have tradeoffs and might be suited for different use cases.
	\end{frame}
	%%%
	
	
%	\begin{frame}%[allowframebreaks]
%		\frametitle{References}
%		\bibliographystyle{amsplain}
%		\bibliography{../assets/bib/refs}
%	\end{frame}
	
\end{document}