% Choose one to switch between slides and handout
%\documentclass[]{beamer}
\documentclass[handout]{beamer}

% Video Meta Data
\title{Bitcoin, Blockchain and Cryptoassets}
\subtitle{CBDC and Stablecoins}
\author{Prof. Dr. Fabian Schär}
\institute{University of Basel}

% Config File
\input{../config/config.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{empty}
\begin{frame}[noframenumbering]
	\titlepage
\end{frame}

%%%

%%%
\begin{frame}{CBDC Background}

Established model of central banks as issuer of physical cash and bank for banks is under pressure.
\vspace{1.5em}	

\textbf{Drivers}
\begin{itemize}
	\item \textbf{Drop in physical cash use} and growing importance of digital payment systems as essential infrastructure.
	\item \textbf{Emergence of new payment solutions} from private sector, with large actors outside the banking industry.
	\item \textbf{Debate on authority and mandate} of central banks around money issuance and payment infrastructure provision.
\end{itemize}

\end{frame}
%%%

%%%

\begin{frame}{CBDC Overview}
CBDC as any \color{focus} digital form of money issued by a central bank. \color{black} 

\begin{center}
	\begin{tikzpicture}[scale=0.4, every node/.style ={scale=0.8}]
		\input{../assets/figures/cube_cbdc}
	\end{tikzpicture}
\end{center}


\uncover<2->{
Selected CBDC design dimensions:
	\begin{itemize}
		\item Centralized vs. decentralized
		\item Retail vs. wholesale
		\item Token-/Object- vs. Account-based
	\end{itemize}
}

\end{frame}
%%%

%%%
\begin{frame}
\frametitle{Title}

\end{frame}
%%%


%%%
\begin{frame}%[allowframebreaks]
\frametitle{References and Recommended Reading}

	\bibliographystyle{amsplain}
	\bibliography{../assets/bib/refs}

\end{frame}
%%%

\end{document}