% Choose one to switch between slides and handout
\documentclass[]{beamer}
%\documentclass[handout]{beamer}

% Video Meta Data
\title{Bitcoin, Blockchain and Cryptoassets}
\subtitle{Introduction to Consensus}
\author{Prof. Dr. Fabian Schär}
\institute{University of Basel}

% Config File
\input{../config/config.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\thispagestyle{empty}
\begin{frame}[noframenumbering]
	\titlepage
\end{frame}

%%%
\begin{frame}{Why Consensus Matters}

Distributed ledger as \color{focus} chain of transactions and states \color{black} whose \color{focus} compliance with an explicit rule set \color{black} is  attested by a reliable network of record keeping nodes.

\uncover<2->{
\vspace{1.5 em}
\textbf{Account statement example:}

\begin{center}
\begin{tikzpicture}[scale=0.7, every node/.style ={scale=0.8}]
		\input{../assets/figures/account_statement_example}
\end{tikzpicture}
\end{center}

\vspace{1 em}

$\Rightarrow$ Value of the ledger content relies on the network attesting it.
}

\end{frame}
%%%	

%%%
\begin{frame}{Challenges for Reaching Consensus}

Two main challenges for consensus in a decentralized network:
	
	\begin{itemize}
		\item<1-> Each participant can adapt the ledger at its own discretion.
		\item<2-> No central authority to decide among two legitimate versions.
	\end{itemize}

\vspace{1 em}
\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style ={scale=0.8}]
		\input{../assets/figures/consensus_challenges}
\end{tikzpicture}
\end{center}

\end{frame}
%%%	

%%%
\begin{frame}{Measures Supporting Consensus}

\textbf{Explicit and unambiguous rule set} for legitimate changes to the ledger and block sequence.
\vspace{0.25 em}

$\Rightarrow$ Invalid blocks are detected easily and unambiguously.
\vspace{1.5 em}

\textbf{Decision mechanism} for consensus over different, legitimate extensions of the ledger. 
\vspace{0.25 em}

$\Rightarrow$ Swiftly resolving situations of uncertainty.
\vspace{1.5 em}

\textbf{Incentive system} that rewards compliant behaviour and / or penalizes manipulation attempts.
\vspace{0.25 em}

$\Rightarrow$ Typically in network currency, tying the participant's interest to the sustainable value of the network.

	
\end{frame}
%%%

%%%
\begin{frame}{What Makes a Good Consensus Mechanism?}

Suitability of a consensus mechanism \color{focus} depends on the purpose and user base \color {black} of a decentralized ledger.
\vspace{1 em}

\textbf{The Trilemma:}

\begin{center}
\begin{tikzpicture}[scale=0.6, every node/.style ={scale=0.8}]
		\input{../assets/figures/trilemma}
\end{tikzpicture}
\end{center}

\textbf{Generalized Rule:} Gains with one goal always come at (some) trade-off to the others.
	
\end{frame}
%%%	


\end{document}